---
title: "Demo distributions"
author: "Math 115 instructional"
date: "2017"
output: html_document
---

```{r include=FALSE,echo=FALSE}

require(tigerstats)

```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

For every distribution there are four commands. The commands for each distribution are prepended with a letter to indicate the functionality:  

“d”	returns the height of the probability density function  
“p”	returns the cumulative density function  
“q”	returns the inverse cumulative density function (quantiles)  
“r”	returns randomly generated numbers  

###We will look at the normal, chi squared, binomial, and t distributions.


# Normal  

###Plot the normal distribution with mean 30 and standard deviation 8 over the interval from 0 to 60.

First create the xaxis to run from 0 to 60

```{r}
xaxis=seq(from=0,to=60,by=.05)
```
Now use dnorm to put in the yaxis values

```{r}
yaxis=dnorm(xaxis,mean=30,sd=8)
```
now plot

```{r}
plot(xaxis,yaxis,type="l")
```
  
  now draw a vertical red line at 42  

```{r}
plot(xaxis,yaxis,type="l")
abline(v=42,col="red")
```


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

#Try now with Chi Squared --- plot a chisq variable with 7 degrees of freedom over the interval from 0 to 60


we can use the same xaxis variable  

we will generate our y axis with dchisq

```{r}
chiyaxis=dchisq(xaxis,df=7)
```
```{r}
plot(xaxis,chiyaxis,type="l")

```

Now suppose we had run a chisq.test on a data-set with 7 degrees of freedom and the value we got for chisq wa 23.

Draw a verticle red line on the distribution function.

```{r}
plot(xaxis,chiyaxis,type="l")
abline(v=23,col="red")
```
  
  what is the area to the left and right of the red line?  
  
  We calculate with pchisq  
 
```{r}
(leftarea=pchisq(q=23,df=7))
(rightarea=pchisq(q=23,df=7,lower.tail = FALSE))
```

Where should we draw the red line to have 95% on its left and 5% on its right?  

We use qchisq for this  

```{r}
point5pct=qchisq(p=.95,df=7)
```
```{r}
plot(xaxis,chiyaxis,type="l",main="five percent above red line")
abline(v=point5pct,col="red")
```

## Challange --- what is the area to the right of the red line you drew 
##at 42 in the first plot??




#The Binomial Distribution

This models drawing a sample of a given size from a population where the probability of a single draw success is known.

plot the binomial distribution of drawing 40 things from a population where a single draw success is 1/6.

```{r}
bxaxis=seq(0:40)
byaxis=dbinom(bxaxis,size=40,prob=1/6)
plot(bxaxis,byaxis,type ="l" )
```

  
  We know that binomial distributions have a mean of p*n (1/6 * 40)
  or a bit less than 7 and a standard deviation of sqrt(p*q*n)  
  about 2.4 --- plot conforms
  
  
  

#Challange  draw a t-distribution with 5 degrees of freedom over the xaxis from -8 to 8.

##Draw a vertical red line at the point where 30 percent of the area is left of your red line.


#Emperical Distribution Functions.

##We dont have the d,p,q,r functionality for data we sample or otherwise we collect.

Here I grab a 1000 point sample from a chisq distribution on 30 deg of freedom

```{r}
mysamp=rchisq(n=1000,df=30)
mydens=ecdf(mysamp)
plot(density(mysamp))
mydens(50)
#so mydens is acting like pnorm for normal 
#what acts like qnorm here?

xx=quantile(mysamp,.95)
abline(v=xx,col="red")
```

